---
title: "Exercise 1"
author: "Emmanuel Ubachi"
date: "April 22, 2023"
output: html_document
---

## Numeric Vectors

### Exercise 1
A vector is a series of values, all of the same type. They are the most basic data type in R and can hold numeric data, character data, or logical data. In R, you can create a vector with the concatenate (or combine) function _c()_. You place the vector elements separated by a comma between the parentheses. For example a numeric vector would look something like this:
_cost <- c(50, 75, 90, 100, 150)_

#### Instructions
Use the function _c()_ to create a numeric vector with the average high temperatures in January for Beijing, Lagos, Paris, Rio de Janeiro, San Juan, and Toronto. The average high temperatures are _35, 88, 42, 84, 81, and 30_ degrees Fahrenheit. Call the object _temp_.
```{r}
# Here is an example creating a numeric vector named cost
cost <- c(50, 75, 90, 100, 150)

# Create a numeric vector to store the temperatures listed in the instructions into a vector named temp
# Make sure to follow the same order in the instructions

city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
temp <- c(35, 88, 42, 84, 81, 30)

names(temp) <- city
temp

```

## Character Vectors

### Exercise 2
As in the previous question, we are going to create a vector. Only this time, we learn to create character vectors. The main difference is that these have to be written as strings and so the names are enclosed within double quotes.

A character vector would look something like this:

_food <- c("pizza", "burgers", "salads", "cheese", "pasta")_

#### instructions
The temperatures we stored in _temp_ are from the cities Beijing, Lagos, Paris, Rio de Janeiro, San Juan, and Toronto.

* Create a vector with these city names and call the object _city_. Make sure to use the correct order, spelling and capitalization!
* (Also note that DataCamp is not compatible with the punctuation used on an iPad or some mobile phones. Complete DataCamp exercises on a full-size computer to avoid correct answer rejections.)

```{r}
# here is an example of how to create a character vector
food <- c("pizza", "burgers", "salads", "cheese", "pasta")

# Create a character vector called city to store the city names
# Make sure to follow the same order as in the instructions

city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

```

## Connecting Numeric and Character Vectors

### Exercise 3
We have successfully assigned the temperatures as numeric values to _temp_ and the city names as character values to _city._ But can we associate the temperature to its related city? Yes! We can do so using a code we already know - _names._ We assign names to the numeric values.

It would look like this:

_cost <- c(50, 75, 90, 100, 150)_
_food <- c("pizza", "burgers", "salads", "cheese", "pasta")_
_names(cost) <- food_

#### Instructions
Use the _names_ function and the objects defined in the previous exercises to associate the temperature data with its corresponding city. (You can go back to the previous questions and copy the objects stored.) Note: to see what happened, after assigning the city names to the temp vector, try printing the temp vector to understand how the names are associated with elements of temp.

```{r}
# Associate the cost values with its corresponding food item
cost <- c(50, 75, 90, 100, 150)
food <- c("pizza", "burgers", "salads", "cheese", "pasta")
names(cost) <- food

# You already wrote this code
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")

# Associate the temperature values with its corresponding city
names(temp) <- city
temp

```

## Subsetting vectors
### Exercise 4

If we want to display only selected values from the object, R can help us do that easily.

For example, if we want to see the cost of the last 3 items in our food list, we would type:

_cost[3:5]_
Note here, that we could also type _cost[c(3,4,5)]_ and get the same result. The _:_ operator helps us condense the code and get consecutive values.

#### Instructions
We will learn to subset using several special operators.

* Use the _[_ and _:_ operators to access the temperature of the first three cities in the list, which are already stored in _temp_.

```{r}
# cost of the last 3 items in our food list:
cost[3:5]

# temperatures of the first three cities in the list:
temp[1:3]

```

### Exercise 5

In the previous question, we accessed the temperature for consecutive cities (1st three). But what if we want to access the temperatures for any 2 specific cities?

An example: To access the cost of _pizza_ (1st) and _pasta_ (5th food item) in our list, the code would be:

_cost[c(1,5)]_

### Instructions
This time we will access our object using just the _[_ operator. Use the _[_ operator to access the temperature of Paris and San Juan already stored in the _temp_ object.

```{r}
# Access the cost of pizza and pasta from our food list 
cost[c(1,5)]

# Define temp
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
names(temp) <- city

# Access the temperatures of Paris and San Juan
temp[c(3,5)]

```

## Sequences
### Exercise 6

The _:_ operator helps us create sequences of numbers. For example, _32:99_ would create a list of numbers from _32_ to _99._

Then, if we want to know the length of this sequence, all we need to do is use the _length_ command.

### Instructions

This time we will use just the _:_ operator. Use the _:_ operator to create a sequence of consecutive integers starting at 12 and ending at 73 and save it in an object _x_, then determine the length of object _x._

```{r}
# Create a vector m of integers that starts at 32 and ends at 99.
m <- 32:99

# Determine the length of object m.
length(m)

# Create a vector x of integers that starts at 12 and ends at 73.
x <- 12:73

# Determine the length of object x.
length(x)

```

### Exersise 7

We can also create different types of sequences in R. For example, in _seq(7, 49, 7)_, the first argument defines the start, and the second the end. The default is to go up in increments of 1, but a third argument lets us tell it by what interval.

#### Instructions
We will learn how to use the _seq()_ function in this question.

8Create a vector containing all the positive odd numbers smaller than 100. The numbers should be in ascending order.

```{r}
# Create a vector with the multiples of 7, smaller than 50.
seq(7, 49, 7) 

# Create a vector containing all the positive odd numbers smaller than 100.
# The numbers should be in ascending order

odd_nums <- seq(1,100,2)

```


## Sequences and length
### Exerceise 8

The second argument of the function _seq_ is actually a maximum, not necessarily the end. So if we type

_seq(7, 50, 7)_

we actually get the same vector of integers as if we type

_seq(7, 49, 7)_

This can be useful because sometimes all we want are sequential numbers that are smaller than some value.

Let's look at an example.

#### Instructions
Create a vector of numbers that starts at 6, does not go beyond 55, and adds numbers in increments of 4/7. So the first three numbers will be 6, 6+4/7, and 6+8/7. How many numbers does the list have? Use only one line of code to answer both questions.

```{r}
# We can create a vector with the multiples of 7, smaller than 50 like this 
seq(7, 49, 7) 

# But note that the second argument does not need to be the last number
# It simply determines the maximum value permitted
# so the following line of code produces the same vector as seq(7, 49, 7)
seq(7, 50, 7)

# Create a sequence of numbers from 6 to 55, with 4/7 increments and determine its length
length(seq(6,55,4/7))

```

##Sequences of certain length
### Exercise 9

The _seq()_ function has another useful argument. The argument length.out. This argument lets us generate sequences that are increasing by the same amount but are of the prespecified length.

For example, this line of code

_x <- seq(0, 100, length.out = 5)_

produces the numbers 0, 25, 50, 75, 100.

Let's create a vector and see what is the class of the object produced.

Instructions

* Determine the class of a vector generated with _seq_ using the length.out argument.
* Specifically, what is the class of the following object _a <- seq(1, 10, length.out = 100)_?

```{r}
# Store the sequence in the object a
a <- seq(1, 10, length.out = 100)

# Determine the class of a
class(a)

```

## Integers
### Exercise 10

We have discussed the numeric class. We just saw that the _seq_ function can generate objects of this class. For another example, type

_class(seq(1, 10, 0.5))_

into the console and note that the _class_ is numeric. R has another type of vector we have not described, the integer class. You can create an integer by adding the letter L after a whole number. If you type

_class(3L)_

in the console, you see this is an integer and not a numeric. For most practical purposes, integers and numerics are indistinguishable. For example 3, the integer, minus 3 the numeric is 0. To see this type this in the console

_3L - 3_

The main difference is that integers occupy less space in the computer memory, so for big computations using integers can have a substantial impact.

#### Instructions
Does this change depending on what we store in an object? What is the class of the following object _a <- seq(1, 10)_?

```{r}
# Store the sequence in the object a
a <- seq(1, 10)

# Determine the class of a
class(a)

```

## Integers and Numerics
### Exercise 11

Let's confirm that _1L_ is an _integer_ not a _numeric_.

#### Instructions

Confirm that the _class_ of _1_ is _numeric_ and the _class_ of _1L_ is _integer_.
```{r}
# Check the class of 1, assigned to the object a
class(1)

# Confirm the class of 1L is integer
class(1L)

```


## Coercion
### Exercise 12

The concept of coercion is a very important one. Watching the video, we learned that when an entry does not match what an R function is expecting, R tries to guess what we meant before throwing an error. This might get confusing at times.

As we've discussed in earlier questions, there are numeric and character vectors. The character vectors are placed in quotes and the numerics are not.

We can avoid issues with coercion in R by changing characters to numerics and vice-versa. This is known as typecasting. The code, _as.numeric(x)_ helps us convert character strings to numbers. There is an equivalent function that converts its argument to a string, _as.character(x)_.

Let's practice doing this!

#### Instructions

* Define the following vector: _x <- c(1, 3, 5,"a")_
* Notice that _x_ has been coerced into a character string.
* If we assign something new to a vector that we've already created, the new definition replaces whatever was previously assigned to the vector.
** Convert _x_ to a vector of numbers using the _as.numeric()_ function. (Note that you will get a warning message, but that is okay!)

```{r}
# Define the vector x
x <- c(1, 3, 5,"a")

# Note that the x is character vector
x

# Redefine `x` to typecast it to get an integer vector using `as.numeric`.
# You will get a warning but that is okay
x <- as.numeric(x)

```















