---
title: "Exercise 1"
author: "Emmanuel Ubachi"
date: "April 22, 2023"
output: html_document
---
## Exercise 1
### str
We're going to be using the following dataset for this module. Run this code in the console.

_library(dslabs)_
_data(murders)_

Next, use the function _str_ to examine the structure of the _murders_ object. We can see that this object is a data frame with 51 rows and five columns.

Which of the following best describes the variables represented in this data frame:
#### Solution
*The state name, the abbreviation of the state name, the state's region, and the state's population and the total number of murders for 2010.


## Exercise 2
### Variable names
In the previous question, we saw the different variables that are a part of this dataset from the output of the _str()_ function. The function _names()_ is specifically designed to extract the column names from a data frame.

#### Instructions
For this question, we will use the _names()_ function to extract the column names of the _murders_ data frame.

*What are the column names used by the _murders_ data frame for the five variables?

```{r}
# Load package and data

library(dslabs)
data(murders)

# Use the function names to extract the variable names

## Solution
names(murders)
```


## Exercise 3
### Examining Variables
In this module we have learned that every variable has a class. For example, the class can be a character, numeric or logical. The function _class()_ can be used to determine the class of an object.

Here we are going to determine the class of one of the variables in the _murders_ data frame. To extract variables from a data frame we use _$_, referred to as the accessor.

In the editor we show an example of how to do this. Let`s try it out for ourselves.

#### Instructions
Use the accessor _$_ to extract the state abbreviations and assign them to the object _a_. What is the class of this object?

```{r}
# To access the population variable from the murders dataset use this code:
p <- murders$population 

# To determine the class of object `p` we use this code:
class(p)

# Use the accessor to extract state abbreviations and assign it to a
a <- murders$abb

# Determine the class of a
class(a)
```

## Exercise 3
### Multiple ways to access variables
An important lesson you should learn early on is that there are multiple ways to do things in R. For example, to generate the first five integers we note that _1:5_ and _seq(1,5)_ return the same result.

There are also multiple ways to access variables in a data frame. For example we can use the square brackets _[[_ instead of the accessor _$_. Example code is included in the editor.

If you instead try to access a column with just one bracket,

_murders["population"]_

R returns a subset of the original data frame containing just this column. This new object will be of class _data.frame_ rather than a vector. To access the column itself you need to use either the _$_ accessor or the double square brackets _[[_.

Parentheses, in contrast, are mainly used alongside functions to indicate what argument the function should be doing something to. For example, when we did _class(p)_ in the last question, we wanted the function _class_ to do something related to the argument _p_.

This is an example of how R can be a bit idiosyncratic sometimes. It is very common to find it confusing at first.

#### Instructions
* Use the square brackets _[[_ to extract the state abbreviations and assign them to the object _b_.
* Then use the _identical_ function to determine if _a_, as defined in the previous exercises, and _b_ are the same.

```{r}
# We extract the population like this:
p <- murders$population

# This is how we do the same with the square brackets:
o <- murders[["population"]] 

# We can confirm these two are the same
identical(o, p)

# Use square brackets to extract `abb` from `murders` and assign it to b
b <- murders[["abb"]]

# Check if `a` and `b` are identical 
identical(a,b)
```


## Exercise 4
### Factors
Using the _str()_ command, we saw that the region column stores a factor. You can corroborate this by using the _class_ command on the region column.

The function _levels_ shows us the categories for the factor. 4

#### Instructions
* For this question we will combine two functions, in a nested way, to determine the number of unique categories. Use the functions _levels_ and _length_ to determine the number of regions defined by this dataset and contained in _murders$region_.

```{r}
# We can see the class of the region variable using class
class(murders$region)

# Determine the number of regions included in this variable
length(levels(murders$region))
```

## Exercise 5
### Tables
The function _table_ takes a vector as input and returns the frequency of each unique element in the vector.

#### Instructions
We will use the table function to answer this question.

* Use the _table_ function in one line of code to create a table showing the number of states per region.

```{r}
# The "c" in `c()` is short for "concatenate," which is the action of connecting items into a chain
# The function `c()` connects all of the strings within it into a single vector, which we can assign to `x`
x <- c("a", "a", "b", "b", "b", "c")
# Here is an example of what the table function does
table(x)

# Write one line of code to show the number of states per region
table(murders$region)

```
